A fast algorithm for global code motion of congruent instructions

### 2017-11-13 cgo/paper.tex
abstract

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
We present a fast \gcm{} (\GCM{}) compiler optimization which schedules congruent
instructions across the program.  Not only \GCM{} saves code size, it exposes
redundancies as well, it exposes more instruction level parallelism in the
basic-block to which instructions are moved, and it enables other passes like
loop invariant motion to remove redundancies. The cost model to drive the code
motion is based on liveness analysis on \SSA{} representation such that the
(virtual) register pressure does not increase resulting in 2\% fewer spills on
the \SPEC{} benchmark suite.

We have implemented the pass in \LLVM{}. It is based on Global Value Numbering
infrastructure available in \LLVM{}. The experimental results show an average
saving of 1\% on the total compilation time on \SPEC{}. \GCM{} enables more
inlining and exposes more loop invariant code motion opportunities in the
majority of benchmarks. We have also seen execution time improvements in a few
of \SPEC{} benchmarks viz. mcf and sjeng, moreover, register spills reduced by
2\% on the \SPEC{} benchmarks suite when compiled for \xlinux{}. \GCM{} is an
optimistic algorithm in the sense that it considers all congruent instructions
in a function as potential candidates. We also formalize why register pressure
reduces as a result of \gcm{}, and how \gcm{} increases instruction level
parallelism thereby enabling more \ooo{} execution on modern architectures.
\end{abstract}

@@

\section{Experimental Evaluation}
\label{sec:experimental-results}
% Experimental results: Register spills, compile time, run time. compile-time comparison w.r.t. gcc
For evaluation of \gcm{}, we built \SPEC{} with the patch. All the experiments
were conducted on an \xlinux{} machine and at -Ofast optimization level. Each
benchmark was run three times and the best result was taken. We collected
execution time and code-size which is listed in Table~\ref{tab:code-size}, other
compiler statistics are listed in Table~\ref{tab:spills}, Table~\ref{tab:stats}, and
Figure~\ref{tab:code-motion-metric}.  We also measured compile time
Table~\ref{tab:compile-time} to see the impact of \gcm{}.

@@

Overall there was only a minor change in the performance numbers with a few improvements 
(429.mcf, 458.sjeng, etc.) and a few regressions (447.dealII, 453.povray,
etc.) The \gcm{} pass was run twice in the pass pipeline (at -Ofast),
first after EarlyCSE pass, and then after the inliner. Because EarlyCSE removes
local redundancies, \GCM{} would have to analyze less redundant instructions, also
as inliner creates more redundancies, \GCM{} would have more opportunities to move
congruent instructions out of branches.

@@

We see some nice gains in code size at -Os with \gcm{} (403.gcc, 447.dealII, and
483.xalancbmk) as shown in Table~\ref{tab:code-size}.  The code size listed here
is the text size delta (from linux size command) of the final executable for
each benchmark.  On the other hand, some benchmarks like 473.astar increased in
code size by almost 2\%. This is because once the code-size of a function
decreases (due to \GCM{}) it becomes cheaper to inline. As we can see in
Table~\ref{tab:stats}, four more functions got inlined in 473.astar. Except for
447.dealII, 450.soplex, 482.sphinx3 and 483.xalancbmk all other benchmarks got
more functions inlined. Also, since the pass runs early, it affects many
optimizations which rely on the number of instructions, length of the use-def
chain, and other metrics.

The code-size at -Ofast is very widespread with huge gains in 483.xalancbmk to
serious regressions in 403.gcc. This is mostly because the compiler's focus is
to improve performance at the cost of code-size at -Ofast, so passes like
inlining, loop-unrolling, code-versioning are aggressive at -Ofast.

\begin{table}[h!]
  \begin{center}
    \scalebox{0.7}{
    \begin{tabular}{|l|r|r|r|}
      \hline
Benchmarks        &base(-Ofast) &\GCM{}(-Ofast) &GCM/base\\\hline
400.perlbench	  & 2539	& 2493	& 0.98 \\\hline
401.bzip2	  & 718	        & 707	& 0.98 \\\hline
403.gcc	          & 5782	& 5722	& 0.99 \\\hline
429.mcf	          & 14	        & 17	& 1.21 \\\hline
433.milc	  & 635	        & 642	& 1.01 \\\hline
444.namd	  & 3223	& 3274	& 1.01 \\\hline
445.gobmk	  & 2166	& 2173	& 1.00 \\\hline
447.dealII	  & 11074	& 10234	& 0.92 \\\hline
450.soplex	  & 1122	& 1125	& 1.00 \\\hline
453.povray	  & 4701	& 4692	& 1.09 \\\hline
456.hmmer	  & 1197	& 1274	& 1.06 \\\hline
458.sjeng	  & 168	        & 176	& 1.05 \\\hline
462.libquantum	  & 118	        & 118	& 1.00 \\\hline
464.h264ref	  & 3379	& 3454	& 1.02 \\\hline
470.lbm	          & 33	        & 33	& 1.00 \\\hline
471.omnetpp	  & 524	        & 527	& 1.00 \\\hline
473.astar	  & 180	        & 201	& 1.12 \\\hline
482.sphinx3	  & 674	        & 670	& 0.99 \\\hline
483.xalancbmk	  & 5102	& 4965	& 0.97 \\\hline
Grand Total	  & 43349	& 42497	& 0.98 \\\hline
    \end{tabular}}
  \end{center}
  \caption{Number of spills with and without\GCM{} on SPEC2006 at
    -Ofast. Lower is better in GCM/base. Spills reduced by 2\%.}
  \label{tab:spills}
\end{table}

\begin{table}[h]
  \begin{center}
    \scalebox{0.7}{
    \begin{tabular}{|l|c|c|c|}
      \hline
      Benchmarks      &   Baseline	        & \GCM{}           & \%Decrease\\
       & (in Millions)& (in Millions)& (lower is better)\\\hline
  400.perlbench	  &   186	& 184	& 0.99 \\\hline
  401.bzip2	  &   76	& 75	& 0.99 \\\hline
  403.gcc         &   1,079	& 1,073	& 0.99 \\\hline
  429.mcf         &   78	& 77	& 0.99 \\\hline
  433.milc	  &   272	& 269	& 0.99 \\\hline
  444.namd	  &   78	& 77	& 0.99 \\\hline
  445.gobmk	  &   261	& 258	& 0.99 \\\hline
  447.dealII	  &   643	& 638	& 0.99 \\\hline
  450.soplex	  &   245	& 242	& 0.99 \\\hline
  453.povray	  &   499	& 494	& 0.99 \\\hline
  456.hmmer	  &   214	& 211	& 0.99 \\\hline
  458.sjeng	  &   89	& 88	& 0.99 \\\hline
  462.libquantum  &   84	& 84	& 0.99 \\\hline
  464.h264ref	  &   151	& 149	& 0.99 \\\hline
  470.lbm         &   71	& 71	& 1.00 \\\hline
  471.omnetpp	  &   382	& 379	& 0.99 \\\hline
  473.astar	  &   78	& 77	& 0.99 \\\hline
  482.sphinx3	  &   158	& 156	& 0.99 \\\hline
  483.xalancbmk	  &   22,485	& 22,455& 1.00 \\\hline
    \end{tabular}}
  \end{center}
  \caption{Change in total number of instructions executed during the
    compilation with and without \GCM{} on \SPEC{} at -Ofast. The
    instructions were counted using valgrind.}
  \label{tab:compile-time}
\end{table}

The compilation time actually reduced for most of the benchmarks as a result of
code motion.  This is because, as \gcm{} removes (congruent) instructions, the
number of instructions to be processed by later compilation passes also
reduces. The compilation time listed here is the total instruction count
executed during the compilation of each benchmark as output from valgrind
-{}-tool=cachegrind \cite{valgrind}.

\begin{table}[h!]
  \begin{center}
    \scalebox{0.8}{
    \begin{tabular}{|l|r|r|r|r|}
      \hline
      &\multicolumn{2}{|c|}{Loop Invariant Motion}  & \multicolumn{2}{|c|}{Functions} \\\hline
 Benchmarks       & Hoisted \% & Sunk \% &  Inlined\%   & Deleted\% \\\hline
 400.perlbench	  & 2.1 & -2.8 &  0.2 &  0.9  \\\hline
 401.bzip2	  & 14.4& 0.0  &  4.7 &  3.1  \\\hline
 403.gcc	  & 13.7& 27.3 &  0.9 &  0.1  \\\hline
 429.mcf	  & -7.3& 0.0  &  0.0 &  0.0  \\\hline
 433.milc	  & 10.2& 0.0  &  8.0 &  0.0  \\\hline
 444.namd	  & 0.1 & 0.0  &  0.8 &  -3.5 \\\hline
 445.gobmk	  & 0.6 & 5.6  &  5.4 &  1.7  \\\hline
 447.dealII	  & 62.2& -86.7& -0.2 &  -0.8 \\\hline
 450.soplex	  & 3.9 & -5.6 & -0.2 &  -0.4 \\\hline
 453.povray	  & -4.8& 39.1 &  0.2 &  -0.1 \\\hline
 456.hmmer	  & 1.0 & 166.7&  0.0 &  -1.9 \\\hline
 458.sjeng	  & 11.5& 0.0  &  6.3 &  0.0  \\\hline
 462.libquantum   & 6.6 & 0.0  &  0.0 &  0.0  \\\hline
 464.h264ref	  & 0.5 & 0.0  &  13.0&  0.0  \\\hline
 470.lbm	  & 0.0 & 0.0  &  0.0 &  0.0  \\\hline
 471.omnetpp	  & 17.0& -6.3 &  1.1 &  -0.4 \\\hline
 473.astar	  & 9.1 & 0.0  &  1.1 &  0.0  \\\hline
 482.sphinx3	  & 5.9 & -18.2&  -2.0&  0.0  \\\hline
 483.xalancbmk	  & 14.6& -3.6 &  -1.0&  -1.2 \\\hline
    \end{tabular}}
  \end{center}
  \caption{Change in the static compile time metrics w.r.t. \GCM{} on
    \SPEC{} at -Ofast. The \% columns show percentage increase in metric
    w.r.t. baseline. \GCM{} improves LICM and inlining in general.}
  \label{tab:stats}
\end{table}

Some useful static metrics are presented in Table~\ref{tab:stats}, which shows
how \gcm{} impacted important compiler optimizations like inlining and
LICM. LICM removes loop invariant code out of the loop. It may hoist the code to
the loop's preheader or sink the code to loop's post-dominator
\cite{steven1997advanced}. As we can see number of instructions hoisted
increased significantly in several benchmarks, although number of instructions
sunk did not change by much.  Number of inlined functions also improved for most
benchmarks. In general when more functions were inlined, more functions were
deleted as in 400.perlbench, 445.gobmk.

The compile time metrics of \GCM{} are listed in
Figure~\ref{tab:code-motion-metric}. The table lists the number of scalars,
loads, stores and calls hoisted as well as removed. For each category, the
number of instructions removed is greater or equal to the number of instructions
hoisted because each code-motion is performed only when at least one identical
computation is found. Loads are hoisted the most followed by scalars, stores and
calls in decreasing order.  This was the common trend in all our
experiments. One reason why loads are hoisted the most is the early execution of
this pass (before mem2reg pass which scalarizes some memory references) in the
\LLVM{} pass pipeline. Passes like mem2reg and instcombine might actually remove
those loads and the number of hoisted loads may change should the \GCM{} pass be
scheduled later. We can see a few sunk instructions even if \LLVM{} has a code
sinking and hoisting transforms in the SimplifyCFG pass which runs before
\GCM{}: this is because sinking and hoisting in SimplifyCFG have some rather
severe limitations to make the analysis and code transform very fast in order to
be able to run the pass several times.  Another reason is that instructions for
which dependencies are not directly available in the successor
(Section~\ref{subsec:legality}) is not sunk for now because that would
require a look ahead on the ability to sink the dependencies. We plan to
implement this in near future.

\section{Conclusions and Future Work}
\label{sec:future-work}

We have presented the \gcm{} of congruent computations in SSA form. We saw that
it improves inlining and LICM opportunities, reduces register spills, it also
improves performance in some cases. To preserve performance and not hoist/sink
too much, we have implemented a register pressure aware cost model as described
in Section~\ref{subsec:cost-models}. A part of our implementation is merged in
LLVM trunk as GVNHoist.cpp and the \GCM{} implementation is under review.

This pass does not hoist fully redundant expressions because they are already
handled by GVN-PRE. It is advisable to run gvn-hoist before and after gvn-pre
because gvn-pre creates opportunities for more instructions to be hoisted.

We would like to integrate \GCM{} with \LLVM{'}s new GVN-PRE implementation such
that more congruent instructions can be scheduled. This would also expose more
redundancies to the GVN-PRE. In general it is a good idea to start with lower
ranked expressions first such that maximum hoisting can happen in one iteration,
however, current implementation does not rank the expressions and iteratively
finds a fixed point when no more candidates are available. Even this
implementation converges quickly and no compile time regression have been
observed. This can be improved by ranking the computations
\cite{rosen1988global}. Also, since \GCM{} runs very early in the pass pipeline,
it will be good to evaluate the code size/performance impact when it is run in
sync with \GVN{}-\PRE{} just like \GCC{} does. With the implementation of \GCM{}
in \LLVM{}, the passes which rely on the code-size or instruction-counts to make
optimization decisions need to be revisited for example, the inliner.

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
We present a fast \gcm{} (\GCM{}) compiler optimization which schedules
congruent instructions across the program. We introduce data structures to help
exploit flow of global values in sparse representation.  Not only \GCM{}
improves performance it saves code size as well, it exposes more instruction
level parallelism in the basic-block to which instructions are moved, and it
enables other passes like loop invariant motion to remove redundancies. The cost
model to drive the code motion is based on liveness analysis on \SSA{}
representation such that the (virtual) register pressure does not increase
resulting in 2\% fewer spills on the \SPEC{} benchmark suite. \GCM{} is an
optimistic algorithm in the sense that it considers all congruent instructions
in a function as potential candidates. We also formalize why register pressure
reduces as a result of \gcm{}, and how \gcm{} increases instruction level
parallelism thereby enabling more \ooo{} execution on modern architectures.

We have implemented the pass in \LLVM{}. The experimental results show
performance gains of 8\% in 462.libquantum and 5.3\% in 471.omnetpp in \SPEC{}
benchmark suite. It also reduces compilation time by reducing number of
instructions to be processed by later compilation stages. \GCM{} enables more
inlining and exposes more loop invariant code motion opportunities in the
majority of benchmarks.
\end{abstract}

@@

\section{Experimental Evaluation}
\label{sec:experimental-results}
% Experimental results: Register spills, compile time, run time. compile-time comparison w.r.t. gcc
For evaluation of \gcm{}, we built \SPEC{} with and without the \GCM{}
optimization. All the experiments were conducted on an \xlinux{} machine and at
-Ofast optimization level with inlining and loop unrolling disabled; inliner and
loop-unroller vary wildly with \GCM{} optimization so it was difficult to
visualize the impact. Each benchmark was run three times and the best result was
taken. We collected execution time and code-size which is listed in
Table~\ref{tab:code-size}, other compiler statistics are listed in
Table~\ref{tab:spills}, Table~\ref{tab:stats}, and
Figure~\ref{tab:code-motion-metric}.  We also measured compile time
Table~\ref{tab:compile-time}. The benchmark 464.h264ref caused an internal
compiler error with the trunk clang so it could not be reported.

@@

Overall, we can see some good improvements in 462.libquantum, 471omnetpp, etc.
and some regressions in 429.mcf, 483.xalancbmk, etc. The \gcm{} pass was run once
in the pass pipeline (at -Ofast), after local common subexpression elimination
(EarlyCSE) pass. Because EarlyCSE removes local redundancies, \GCM{} would have
to analyze less redundant instructions. We believe some performance regressions
can be mitigated by running the \GCM{} few more times.

We see some code size reduction in 400.perlbench, 403.gcc, and 453.povray as
shown in Table~\ref{tab:code-size}.  The code size listed here is the text size
delta (from size command in linux) of the final executable for each benchmark.  On
the other hand 447.dealII increased in code size. This is because once the
code-size of a function decreases (due to \GCM{}) it becomes cheaper to do some
optimizations like inlining, loop unrolling, copy propagation, rematerialization
etc. We tried to minimize the noise by disabling inliner, and loop unroller but
still we can see some increase in the code size.  To see the impact of \GCM{} on
inliner we ran \GCM{} with inliner enabled and collected the compile-time
statistics produced by LLVM (See Table~\ref{tab:stats}) . Except for 447.dealII,
450.soplex, 482.sphinx3 and 483.xalancbmk all other benchmarks got more
functions inlined. Also, since the pass runs early, it affects many
optimizations which rely on the number of instructions, length of the use-def
chain, and other metrics.

\begin{table}[h!]
  \begin{center}
    \scalebox{0.8}{
    \begin{tabular}{|l|r|r|r|r|}
      \hline
      &\multicolumn{2}{|c|}{Loop Invariant Motion}  & \multicolumn{2}{|c|}{Functions} \\\hline
 Benchmarks       & Hoisted\% & Sunk\% &  Inlined\%   & Deleted\% \\\hline
 400.perlbench	  & 2.1 & -2.8 &  0.2 &  0.9  \\\hline
 401.bzip2	  & 14.4& 0.0  &  4.7 &  3.1  \\\hline
 403.gcc	  & 13.7& 27.3 &  0.9 &  0.1  \\\hline
 429.mcf	  & -7.3& 0.0  &  0.0 &  0.0  \\\hline
 433.milc	  & 10.2& 0.0  &  8.0 &  0.0  \\\hline
 444.namd	  & 0.1 & 0.0  &  0.8 &  -3.5 \\\hline
 445.gobmk	  & 0.6 & 5.6  &  5.4 &  1.7  \\\hline
 447.dealII	  & 62.2& -86.7& -0.2 &  -0.8 \\\hline
 450.soplex	  & 3.9 & -5.6 & -0.2 &  -0.4 \\\hline
 453.povray	  & -4.8& 39.1 &  0.2 &  -0.1 \\\hline
 456.hmmer	  & 1.0 & 166.7&  0.0 &  -1.9 \\\hline
 458.sjeng	  & 11.5& 0.0  &  6.3 &  0.0  \\\hline
 462.libquantum   & 6.6 & 0.0  &  0.0 &  0.0  \\\hline
 470.lbm	  & 0.0 & 0.0  &  0.0 &  0.0  \\\hline
 471.omnetpp	  & 17.0& -6.3 &  1.1 &  -0.4 \\\hline
 473.astar	  & 9.1 & 0.0  &  1.1 &  0.0  \\\hline
 482.sphinx3	  & 5.9 & -18.2&  -2.0&  0.0  \\\hline
 483.xalancbmk	  & 14.6& -3.6 &  -1.0&  -1.2 \\\hline
    \end{tabular}}
  \end{center}
  \caption{Change in the static compile time metrics w.r.t. \GCM{} on \SPEC{} at
    -Ofast. The \% columns show percentage increase in metric w.r.t. baseline,
    i.e., without \GCM{}. \GCM{} improves LICM and inlining in general.}
  \label{tab:stats}
\end{table}

Table~\ref{tab:stats}, which shows how \gcm{} impacted important compiler
optimizations like inlining and loop invariant code motion (LICM). LICM removes
loop invariant code out of the loop. It may hoist the code to the loop's
preheader or sink the code to loop's post-dominator
\cite{steven1997advanced}. As we can see number of instructions hoisted
increased significantly in several benchmarks, although number of instructions
sunk did not change by much.  Number of inlined functions also improved for most
benchmarks. In general when more functions were inlined, more functions were
deleted as in 401.bzip2, 445.gobmk.

\begin{table}[h!]
  \begin{center}
    \scalebox{0.7}{
    \begin{tabular}{|l|r|r|r|}
      \hline
Benchmarks        &base(-Ofast) &\GCM{}(-Ofast) &GCM/base\\\hline
400.perlbench	  & 2539	& 2493	& 0.98 \\\hline
401.bzip2	  & 718	        & 707	& 0.98 \\\hline
403.gcc	          & 5782	& 5722	& 0.99 \\\hline
429.mcf	          & 14	        & 17	& 1.21 \\\hline
433.milc	  & 635	        & 642	& 1.01 \\\hline
444.namd	  & 3223	& 3274	& 1.01 \\\hline
445.gobmk	  & 2166	& 2173	& 1.00 \\\hline
447.dealII	  & 11074	& 10234	& 0.92 \\\hline
450.soplex	  & 1122	& 1125	& 1.00 \\\hline
453.povray	  & 4701	& 4692	& 1.09 \\\hline
456.hmmer	  & 1197	& 1274	& 1.06 \\\hline
458.sjeng	  & 168	        & 176	& 1.05 \\\hline
462.libquantum	  & 118	        & 118	& 1.00 \\\hline
470.lbm	          & 33	        & 33	& 1.00 \\\hline
471.omnetpp	  & 524	        & 527	& 1.00 \\\hline
473.astar	  & 180	        & 201	& 1.12 \\\hline
482.sphinx3	  & 674	        & 670	& 0.99 \\\hline
483.xalancbmk	  & 5102	& 4965	& 0.97 \\\hline
Grand Total	  & 43349	& 42497	& 0.98 \\\hline
    \end{tabular}}
  \end{center}
  \caption{Number of spills with and without\GCM{} on SPEC2006 at
    -Ofast. Lower is better in GCM/base. Spills reduced by 2\%.}
  \label{tab:spills}
\end{table}

\begin{table}[h]
  \begin{center}
    \scalebox{0.7}{
    \begin{tabular}{|l|c|c|c|}
      \hline
      Benchmarks      &   Baseline	        & \GCM{}           & \%Decrease\\
       & (in Millions)& (in Millions)& (lower is better)\\\hline
  400.perlbench	  &   186	& 184	& 0.99 \\\hline
  401.bzip2	  &   76	& 75	& 0.99 \\\hline
  403.gcc         &   1,079	& 1,073	& 0.99 \\\hline
  429.mcf         &   78	& 77	& 0.99 \\\hline
  433.milc	  &   272	& 269	& 0.99 \\\hline
  444.namd	  &   78	& 77	& 0.99 \\\hline
  445.gobmk	  &   261	& 258	& 0.99 \\\hline
  447.dealII	  &   643	& 638	& 0.99 \\\hline
  450.soplex	  &   245	& 242	& 0.99 \\\hline
  453.povray	  &   499	& 494	& 0.99 \\\hline
  456.hmmer	  &   214	& 211	& 0.99 \\\hline
  458.sjeng	  &   89	& 88	& 0.99 \\\hline
  462.libquantum  &   84	& 84	& 0.99 \\\hline
  470.lbm         &   71	& 71	& 1.00 \\\hline
  471.omnetpp	  &   382	& 379	& 0.99 \\\hline
  473.astar	  &   78	& 77	& 0.99 \\\hline
  482.sphinx3	  &   158	& 156	& 0.99 \\\hline
  483.xalancbmk	  &   22,485	& 22,455& 1.00 \\\hline
    \end{tabular}}
  \end{center}
  \caption{Change in total number of instructions executed during the
    compilation with and without \GCM{} on \SPEC{} at -Ofast. The
    instructions were counted using valgrind.}
  \label{tab:compile-time}
\end{table}

The compilation time actually reduced for most of the benchmarks as a result of
code motion.  This is because, as \gcm{} removes (congruent) instructions, the
number of instructions to be processed by later compilation passes also
reduces. The compilation time listed here is the total instruction count
executed during the compilation of each benchmark as output from valgrind
-{}-tool=cachegrind \cite{valgrind}. We used valgrind because it was a simple
and reliable way to show the impact of our optimization.

@@

The compile time metrics of \GCM{} are listed in
Figure~\ref{tab:code-motion-metric}. The table lists the number of scalars,
loads, stores and calls hoisted as well as removed. For each category, the
number of instructions removed is greater or equal to the number of instructions
hoisted because each code-motion is performed only when at least one identical
computation is found. Loads are hoisted the most followed by scalars, stores and
calls in decreasing order.  This was the common trend in all our
experiments. One reason why loads are hoisted the most is the early execution of
this pass (before mem2reg pass which scalarizes some memory references) in the
\LLVM{} pass pipeline. Passes like mem2reg and instcombine might actually remove
those loads and the number of hoisted loads may change should the \GCM{} pass be
scheduled later. We can see only a few sunk instructions because hoisting is
performed before sinking (Section~\ref{subsec:optimistic}) and some of the
candidates which can either be hoisted or sunk got hoisted.

\section{Conclusions and Future Work}
\label{sec:future-work}
We have presented a fast algorithm to perform \gcm{} of congruent computations
in SSA form. We introduced data structures like $\chi$ nodes, $\Phi$ nodes which
helps exploit flow of values in sparse representation.  Experimental results
show that it is a useful performance and code-size optimization. It also enables
inlining and LICM opportunities, reduces register spill. To preserve performance
and not hoist/sink too much, we have implemented a register pressure aware cost
model as described in Section~\ref{subsec:cost-models}. A part of our
implementation is merged in LLVM trunk as GVNHoist.cpp and the \GCM{}
implementation is under review. This pass does not hoist fully redundant
expressions because they are already handled by a later partial redundancy
elimination pass.

We would like to explore other places in the optimization pipeline where \GCM{}
could improve performance. It may be useful to integrate \GCM{} with \LLVM{'}s
new GVN-PRE implementation such that more congruent instructions can be
scheduled. This would also expose more redundancies to the GVN-PRE. Also, since
\GCM{} runs very early in the pass pipeline, it will be good to evaluate the
code size/performance impact when it is run in sync with \GVN{}-\PRE{} just like
\GCC{} does. With the implementation of \GCM{} in \LLVM{}, the passes which rely
on the code-size or instruction-counts to make optimization decisions need to be
revisited for example, the inliner, the loop unroller etc.
